The Changing Shape of Empire: A GM Study of Chimú Bottles in Museum Collections

### 2022-02-05 supp/03-gm.botjar.Rmd
<init .s>
<edit index>
<init plots>

# Geometric morphometrics

@@

```{r pca.warp, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap="Results of PCA summarising shape variation in the aggregated sample; gray, north-formative/early; orange, north-late/historic; brown, south-formative/early; and black, south-late/historic; with mean shapes (gray) contrasted with max/min shapes (black)."}
# pca warp                   
#knitr::include_graphics('images/pca-warp.jpg')
```

## Vessel allometry

```{r allometry.1, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# allometry: shape as a function of vessel size?
fit.size<-procD.lm(shape ~ size, 
                   data = gdf, 
                   print.progress = FALSE, 
                   iter = 9999)

# Procrustes ANOVA: does shape change with size?
anova(fit.size)
```

```{r common.allometry.1, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# common allometries
fit.common<-procD.lm(shape ~ size + type, 
                     data = gdf, 
                     print.progress = FALSE, 
                     iter = 9999)

anova(fit.common)

common.allometry <- plot(fit.size,
                         type = "regression",
                         reg.type = "RegScore",
                         predictor = log(gdf$size),
                         pch = pch.gps.type,
                         col = col.gps.type)

# common allometric component (Mitteroecker 2004)
plotAllometry(fit.size, 
              size = gdf$size, 
              logsz = TRUE, 
              method = "CAC", 
              pch = pch.gps.type, 
              col = col.gps.type)
```

```{r unique.allometry.1, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# unique allometries
fit.unique<-procD.lm(shape ~ size * type, 
                     data = gdf, 
                     print.progress = FALSE, 
                     iter = 9999)

anova(fit.unique)

# do vessel forms express parallel, convergent, or divergent morphological characteristics?
plotAllometry(fit.unique, 
              size = gdf$size, 
              logsz = TRUE, 
              method = "PredLine", 
              pch = pch.gps.type, 
              col = col.gps.type)
```

>>>>

# Bottle vs Jar morphology

@@

### Minima/maxima of PC1/2 with warp grids

```{r min.max.botjar, echo=TRUE, out.width = "100%", dpi = 300, warning=FALSE}
# plot x/y maxima/minima
## x - minima
ref <- mean.shape <- mshape(Y.gpa$coords)
plotRefToTarget(M1 = ref, 
                M2 = pca$shapes$shapes.comp1$min, 
                method = "TPS")

plotRefToTarget(M1 = ref, 
                M2 = pca$shapes$shapes.comp1$max, 
                method = "TPS")

plotRefToTarget(M1 = ref, 
                M2 = pca$shapes$shapes.comp2$min, 
                method = "TPS")

plotRefToTarget(M1 = ref, 
                M2 = pca$shapes$shapes.comp2$max,
                method = "TPS")
```

```{r pca.warp.botjar, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap="Results of PCA summarising shape variation in bottles, gray; and jars, orange; with shapes + warp grids for maxima/minima of each axis."}
# pca warp                   
knitr::include_graphics('images/pca-warp-botjar.jpg')
```

### 2022-02-05 supp/04-gm.neck.Rmd
<init plots>

```{r load.data.s, echo = TRUE}
# load packages
#devtools::install_github("mlcollyer/RRPP")
#devtools::install_github("geomorphR/geomorph", ref = "Stable", build_vignettes = TRUE)
library(here)
library(geomorph)
library(tidyverse)
library(wesanderson)

@@

# palette
pal <- wes_palette("Moonrise2", 2, type = "continuous")

@@

```{r pca.warp.s, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap="Results of PCA summarising shape variation in the aggregated sample; gray, north-formative/early; orange, north-late/historic; brown, south-formative/early; and black, south-late/historic; with mean shapes (gray) contrasted with max/min shapes (black)."}
# pca warp                   
#knitr::include_graphics('images/pca-warp.jpg')
```

@@

```{r mshape.s, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# subset landmark coordinates to produce mean shapes
new.coords<-coords.subset(A = Y.gpa$coords,
                          group = qdata$type)
names(new.coords)

>>>>

```{r load.data.s, echo = TRUE}
# load packages
library(here)
library(geomorph)
library(tidyverse)
library(wesanderson)

@@

# palette
pal <- wes_palette("Moonrise2")

@@

### Minima/maxima of PC1/2 with warp grids

```{r min.max.neck, echo=TRUE, out.width = "100%", dpi = 300, warning=FALSE}
# plot x/y maxima/minima
## x - minima
ref <- mean.shape <- mshape(Y.gpa$coords)
plotRefToTarget(M1 = ref, 
                M2 = pca$shapes$shapes.comp1$min, 
                method = "TPS")

plotRefToTarget(M1 = ref, 
                M2 = pca$shapes$shapes.comp1$max, 
                method = "TPS")

plotRefToTarget(M1 = ref, 
                M2 = pca$shapes$shapes.comp2$min, 
                method = "TPS")

plotRefToTarget(M1 = ref, 
                M2 = pca$shapes$shapes.comp2$max,
                method = "TPS")
```

```{r pca.warp.ne, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap="Results of PCA summarising shape variation in bottle necks, gray; and jar necks, orange; with shapes + warp grids for maxima/minima of each axis."}
# pca warp neck                   
knitr::include_graphics('images/pca-warp-neck.jpg')
```

@@

```{r mshape.s, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap="Mean shapes for bottles (a), jars (b), and comparison of bottle (gray) and jar (black) neck shapes (c)."}
# subset landmark coordinates to produce mean shapes
new.coords<-coords.subset(A = Y.gpa$coords,
                          group = qdata$type)
names(new.coords)

### 2022-02-05 supp/05-gm.body.Rmd
<init plots>

```{r load.data.b, echo = TRUE}
# load packages
#devtools::install_github("mlcollyer/RRPP")
#devtools::install_github("geomorphR/geomorph", ref = "Stable", build_vignettes = TRUE)
library(here)
library(geomorph)
library(tidyverse)
library(wesanderson)

@@

```{r pca.warp.b, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap="Results of PCA summarising shape variation in the aggregated sample; gray, north-formative/early; orange, north-late/historic; brown, south-formative/early; and black, south-late/historic; with mean shapes (gray) contrasted with max/min shapes (black)."}
# pca warp                   
#knitr::include_graphics('images/pca-warp.jpg')
```

@@

## Mean shapes

```{r mshape.b, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# subset landmark coordinates to produce mean shapes
new.coords<-coords.subset(A = Y.gpa$coords,
                          group = qdata$type)
names(new.coords)

# type shape means
mean <- lapply(new.coords, mshape)

# plot mean shapes (type)
plot(mean$bottle)
plot(mean$jar)

# comparison plots
plotRefToTarget(mean$bottle,
                mean$jar, 
                method = c("points"),
                mag = 1)

#knitr::include_graphics('images/mshape-body.png')
```

>>>>

```{r load.data.b, echo = TRUE}
# load packages
library(here)
library(geomorph)
library(tidyverse)
library(wesanderson)

@@

### Minima/maxima of PC1/2 with warp grids

```{r min.max.body, echo=TRUE, out.width = "100%", dpi = 300, warning=FALSE}
# plot x/y maxima/minima
## x - minima
ref <- mean.shape <- mshape(Y.gpa$coords)
plotRefToTarget(M1 = ref, 
                M2 = pca$shapes$shapes.comp1$min, 
                method = "TPS")

plotRefToTarget(M1 = ref, 
                M2 = pca$shapes$shapes.comp1$max, 
                method = "TPS")

plotRefToTarget(M1 = ref, 
                M2 = pca$shapes$shapes.comp2$min, 
                method = "TPS")

plotRefToTarget(M1 = ref, 
                M2 = pca$shapes$shapes.comp2$max,
                method = "TPS")
```

```{r pca.warp.bo, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap="Results of PCA summarising shape variation in bottle bodies, gray; and jar bodies, orange; with shapes + warp grids for maxima/minima of each axis."}
# pca warp body                 
knitr::include_graphics('images/pca-warp-body.jpg')
```

@@

### 2022-02-05 supp/index.Rmd
<init plots>

```{r botjar, out.width = "100%", dpi = 300, echo=FALSE, warning=FALSE, fig.cap="Mean shapes for bottles (a), jars (c), and comparison of bottle (gray) and jar (black) shape (b)."}
knitr::include_graphics('images/mshape.png')
```

>>>>

```{r pca.warp.botjar.index, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap="Results of PCA summarising shape variation in bottles, gray; and jars, orange; with shapes + warp grids for maxima/minima of each axis."}
# pca warp neck                   
knitr::include_graphics('images/pca-warp-botjar.jpg')
``` 
```{r botjar, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap="Mean shapes for bottles (a), jars (c), and comparison of bottle (gray) and jar (black) shape (b)."}
knitr::include_graphics('images/mshape.png')
```

**Neck morphology**:

 * Bottle and jar necks **differ significantly** in shape
 * Bottle and jar necks do not differ in size

```{r pca.warp.ne.index, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap="Results of PCA summarising shape variation in bottle necks, gray; and jar necks, orange; with shapes + warp grids for maxima/minima of each axis."}
# pca warp neck                   
knitr::include_graphics('images/pca-warp-neck.jpg')
```
```{r botjarneck, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap="Mean shapes for bottle necks (a), jar necks (b), and comparison of bottle (gray) and jar (black) neck shapes (c)."}
knitr::include_graphics('images/mshape-neck.png')
```

**Basal morphology**:

 * Bottle and jar bases **differ significantly** in shape
 * Bottle and jar bases do not differ in size

```{r pca.warp.ba.index, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap="Results of PCA summarising shape variation in bottle bases, gray; and jar bases, orange; with shapes + warp grids for maxima/minima of each axis."}
# pca warp                   
knitr::include_graphics('images/pca-warp-base.jpg')
```
```{r botjarbase.index, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap="Mean shapes for bottle bases (a), jar bases (b), and comparison of bottle (gray) and jar (black) base shapes (c)."}
knitr::include_graphics('images/mshape-base.png')
```
